Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-02-07 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\begin{abstract} 
%CAUSE --> EFFECT --> MOTIVATION (P1)
A critical aspect of multiphase flows in porous media is the displacement efficiency that measures the quantity of fluid that can be pushed by another fluid due to pressure gradient. Migration of contaminants and waterflooding/EOR (enhanced oil recovery) are typical applications in which simulation of immiscible fluid displacement dynamics may help mitigating water resources contamination and improving hydrocarbons production. In this work, flow instabilities that arise from fluids' viscosity ratio are numerically investigated. Viscous flow instabilities lead to formation and growth of fingers that have strong impact on the efficiency of oil and gas exploration.

%CAUSE --> EFFECT --> GAPS IN KNOWLEDGE --> AIMS OF THE WORK (P2)
From initial fluid flow studies of the Hele-Shaw experimental apparatus, a mathematically analogous process to the two-dimensional flow in porous media problem, fingers development as a function of mobility ratio (MR), different domain set-up under uniform or different permeabilities regions and mesh types is examined. A series of cases are listed, showing the validation of the method against experimental setup followed by a more complex 2D domain setup under different viscosity ratios and different mesh resolutions. Finally, a 3D domain setup is showcasing the difference between the preferential flow path and the fingering regimes that will occur in the domain.

>>>>

\begin{abstract} 
%CAUSE --> EFFECT --> MOTIVATION (P1)
  A critical aspect of multiphase flows in porous media is the displacement efficiency that measures the quantity of fluid that can be pushed by another fluid due to pressure gradient. Migration of contaminants and waterflooding/EOR (enhanced oil recovery) are typical applications in which simulation of immiscible fluid displacement dynamics may help mitigating water resources contamination and improving hydrocarbons production. Due to large fluids' viscosity ratio, flow instabilities may arise leading to formation and growth of fingers with strong impact on the efficiency of oil and gas exploration.

%CAUSE --> EFFECT --> GAPS IN KNOWLEDGE --> AIMS OF THE WORK (P2)
The main aim of this work is to numerically investigate

  In this work, flow instabilities that arise from fluids' viscosity ratio are numerically investigated.

From initial fluid flow studies of the Hele-Shaw experimental apparatus, a mathematically analogous process to the two-dimensional flow in porous media problem, fingers development as a function of mobility ratio (MR), different domain set-up under uniform or different permeabilities regions and mesh types is examined. A series of cases are listed, showing the validation of the method against experimental setup followed by a more complex 2D domain setup under different viscosity ratios and different mesh resolutions. Finally, a 3D domain setup is showcasing the difference between the preferential flow path and the fingering regimes that will occur in the domain.
